---
title: "Descriptive Analysis – Real Estate Properties"
author: "Vicente Emilio Tralci Sindoni"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5)
```

# Package Setup

```{r packages}
pkgs <- c("tidyverse", "skimr", "corrplot", "clickR",
          "ggplot2", "scales", "knitr", "kableExtra", "effectsize")
for (p in pkgs) if (!require(p, character.only = TRUE)) install.packages(p)

library(tidyverse); library(skimr); library(corrplot); library(clickR)
library(ggplot2);   library(scales); library(knitr);   library(kableExtra)
library(effectsize)
```

# Load Data

```{r load-data}
df <- read_csv("/Users/vets/projects/upv/1.2/pr1/proyecto1_IIA/data/trainData.csv")

cat("Rows:", nrow(df), "\n")
cat("Columns:", ncol(df), "\n")
glimpse(df)
```

```{r derive-geo-vars}
# ── French département/région maps ────────────────────────────────────────────
REGION_MAP <- list(
  "Auvergne-Rhône-Alpes"          = c("01","03","07","15","26","38","42","43","63","69","73","74"),
  "Bourgogne-Franche-Comté"       = c("21","25","39","58","70","71","89","90"),
  "Bretagne"                      = c("22","29","35","56"),
  "Centre-Val de Loire"           = c("18","28","36","37","41","45"),
  "Corse"                         = c("20"),
  "Grand Est"                     = c("08","10","51","52","54","55","57","67","68","88"),
  "Hauts-de-France"               = c("02","59","60","62","80"),
  "Île-de-France"                 = c("75","77","78","91","92","93","94","95"),
  "Normandie"                     = c("14","27","50","61","76"),
  "Nouvelle-Aquitaine"            = c("16","17","19","23","24","33","40","47","64","79","86","87"),
  "Occitanie"                     = c("09","11","12","30","31","32","34","46","48","65","66","81","82"),
  "Pays de la Loire"              = c("44","49","53","72","85"),
  "Provence-Alpes-Côte d'Azur"   = c("04","05","06","13","83","84")
)

# Flatten to named vector: dept_code → region_name
region_lookup <- setNames(
  rep(names(REGION_MAP), lengths(REGION_MAP)),
  unlist(REGION_MAP)
)

# Derive department and region on df itself (available to all downstream chunks)
df <- df %>%
  mutate(
    department = str_sub(as.character(postal_code), 1, 2),
    department = if_else(str_to_upper(department) %in% c("2A","2B"),
                         str_to_upper(department), department),
    region     = region_lookup[department],
    region     = if_else(is.na(region), "unknown", region)
  )

cat("department — unique values:", n_distinct(df$department, na.rm = TRUE), "\n")
cat("region     — unique values:", n_distinct(df$region,     na.rm = TRUE), "\n")
```
---

# Dataset Overview

## Full Skimr Summary

```{r skim}
skim(df)
```

## clickR Descriptive

```{r descriptive}
descriptive(df)
```

---

# Variable Types

```{r var-types}
var_types <- data.frame(
  Variable  = names(df),
  Type      = sapply(df, function(x) class(x)),
  row.names = NULL
)

var_types %>%
  kable(caption = "Variable types in the dataset") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

---

# Unique Value Counts

```{r unique-counts}
unique_counts <- df %>%
  summarise(across(everything(), ~ n_distinct(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Unique_Values") %>%
  arrange(desc(Unique_Values))

unique_counts %>%
  kable(caption = "Number of unique values per variable") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

---

# Categorical Variables

## Count per Category

```{r cat-vars}
cat_vars      <- df %>% select(where(~ is.character(.x) | is.factor(.x)))
cat_var_names <- names(cat_vars)

if (length(cat_var_names) == 0) cat("No categorical variables detected.\n") else
  cat("Categorical variables:", paste(cat_var_names, collapse = ", "), "\n")
```

```{r cat-freq-tables}
for (var in cat_var_names) {
  cat("\n\n### Frequency table:", var, "\n\n")
  freq_tbl <- df %>%
    count(.data[[var]], name = "Count") %>%
    arrange(desc(Count)) %>%
    mutate(Proportion = percent(Count / sum(Count), accuracy = 0.1))
  print(
    freq_tbl %>%
      kable(caption = paste("Frequency table:", var)) %>%
      kable_styling(bootstrap_options = c("striped","hover"), full_width = FALSE)
  )
}
```

## Bar Charts per Categorical Variable

```{r cat-bar-charts, fig.height=5}
for (var in cat_var_names) {
  freq_data <- df %>%
    count(.data[[var]], name = "Count") %>%
    arrange(desc(Count)) %>%
    mutate(!!var := fct_reorder(.data[[var]], Count, .desc = TRUE))

  print(arrange(freq_data, Count), n = 100)

  p <- ggplot(freq_data, aes(x = .data[[var]], y = Count, fill = Count)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = Count), vjust = -0.4, size = 3.2) +
    scale_fill_gradient(low = "#a8c8e8", high = "#1a5fa8") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
    labs(title = paste("Count by", var), x = var, y = "Count") +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  print(p)
}
```

---

# Numerical Variables

## Summary Statistics

```{r num-summary}
num_vars <- df %>% select(where(is.numeric))

num_summary <- num_vars %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Value") %>%
  group_by(Variable) %>%
  summarise(
    N      = sum(!is.na(Value)),
    Mean   = round(mean(Value,   na.rm = TRUE), 2),
    Median = round(median(Value, na.rm = TRUE), 2),
    SD     = round(sd(Value,     na.rm = TRUE), 2),
    Min    = round(min(Value,    na.rm = TRUE), 2),
    Max    = round(max(Value,    na.rm = TRUE), 2),
    IQR    = round(IQR(Value,    na.rm = TRUE), 2)
  ) %>%
  arrange(Variable)

num_summary %>%
  kable(caption = "Numerical variable summary statistics") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

---

# Missing Values

## Missing Count and Percentage per Variable

```{r missing-values}
missing_df <- df %>%
  summarise(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "Variable", values_to = "Missing") %>%
  mutate(Total = nrow(df), Pct_Missing = round(Missing / Total * 100, 2)) %>%
  arrange(desc(Missing))          # ← removed filter(Missing > 0)

if (nrow(missing_df) == 0) cat("No missing values found.\n") else
  missing_df %>%
    kable(caption = "Missing values per variable") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Missing Values Bar Chart

```{r missing-chart}
if (nrow(missing_df) > 0) {
  ggplot(
    missing_df %>% mutate(Variable = fct_reorder(Variable, Pct_Missing, .desc = TRUE)),
    aes(x = Variable, y = Pct_Missing, fill = Pct_Missing)
  ) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = paste0(Pct_Missing, "%")), vjust = -0.4, size = 3.2) +
    scale_fill_gradient(low = "#f7c59f", high = "#e05c1a") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)),
                       labels = function(x) paste0(x, "%")) +
    labs(title = "Percentage of Missing Values by Variable",
         x = "Variable", y = "% Missing") +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

---

# Price Distribution

```{r price-dist}
price_col <- "price"

if (price_col %in% names(df)) {
  price_data <- df %>% filter(!is.na(.data[[price_col]]))
  med_val    <- median(price_data[[price_col]], na.rm = TRUE)
  mean_val   <- mean(price_data[[price_col]],   na.rm = TRUE)

  ggplot(price_data, aes(x = .data[[price_col]])) +
    geom_histogram(bins = 50, fill = "#1a5fa8", colour = "white", alpha = 0.85) +
    geom_vline(xintercept = med_val,  colour = "#e05c1a", linetype = "dashed", linewidth = 0.9) +
    geom_vline(xintercept = mean_val, colour = "#2ca02c", linetype = "dashed", linewidth = 0.9) +
    annotate("text", x = med_val,  y = Inf,
             label = paste("Median:", comma(med_val)),
             vjust = 1.5, hjust = -0.05, colour = "#e05c1a", size = 3.5) +
    annotate("text", x = mean_val, y = Inf,
             label = paste("Mean:", comma(mean_val)),
             vjust = 3, hjust = -0.05, colour = "#2ca02c", size = 3.5) +
    scale_x_continuous(labels = comma) +
    labs(title = paste("Distribution of", price_col), x = price_col, y = "Count") +
    theme_minimal(base_size = 12)
}
```

```{r price-boxplot}
if (price_col %in% names(df)) {
  ggplot(df %>% filter(!is.na(.data[[price_col]])),
         aes(y = .data[[price_col]])) +
    geom_boxplot(fill = "#a8c8e8", colour = "#1a5fa8",
                 outlier.colour = "#e05c1a", outlier.alpha = 0.4, width = 0.4) +
    scale_y_continuous(labels = comma) +
    labs(title = paste("Boxplot of", price_col), y = price_col) +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
}
```

```{r price-log}
if (price_col %in% names(df)) {
  ggplot(df %>% filter(!is.na(.data[[price_col]]), .data[[price_col]] > 0),
         aes(x = .data[[price_col]])) +
    geom_histogram(bins = 50, fill = "#6a3fa8", colour = "white", alpha = 0.85) +
    scale_x_log10(labels = comma) +
    labs(title = paste("Distribution of", price_col, "(log10 scale)"),
         x = paste0("log10(", price_col, ")"), y = "Count") +
    theme_minimal(base_size = 12)
}
```

---

# City Analysis

```{r city-config}
city_col  <- "city"
top_n_val <- 20
```

## Top Cities by Listings

```{r city-listings}
if (city_col %in% names(df)) {
  city_counts <- df %>%
    count(.data[[city_col]], name = "Listings") %>%
    arrange(desc(Listings)) %>%
    slice_head(n = top_n_val) %>%
    mutate(!!city_col := fct_reorder(.data[[city_col]], Listings, .desc = TRUE))

  ggplot(city_counts, aes(x = .data[[city_col]], y = Listings, fill = Listings)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = comma(Listings)), vjust = -0.4, size = 3) +
    scale_fill_gradient(low = "#a8c8e8", high = "#1a5fa8") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)), labels = comma) +
    labs(title = paste("Top", top_n_val, "Cities by Number of Listings"),
         x = "City", y = "Listings") +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```
```{r fn-cities-per-dept}
# ── Reusable helper ───────────────────────────────────────────────────────────
# Returns a tibble with department, listing count, and cities.
# Args:
#   data        : the main data frame
#   postal_col  : name of the postal code column (string)
#   city_col    : name of the city column (string)
#   min_listings: if set, filters to departments with FEWER than this value
# Usage examples:
#   cities_per_dept(df)
#   cities_per_dept(df, min_listings = 30)
#   cities_per_dept(df, postal_col = "zip", city_col = "municipality")

cities_per_dept <- function(data,
                            postal_col   = "postal_code",
                            city_col     = "city",
                            min_listings = NULL) {

  # Validate columns
  missing_cols <- setdiff(c(postal_col, city_col), names(data))
  if (length(missing_cols) > 0) {
    stop("Column(s) not found in data: ", paste(missing_cols, collapse = ", "))
  }

  result <- data %>%
    mutate(department = str_sub(as.character(.data[[postal_col]]), 1, 2)) %>%
    group_by(department) %>%
    summarise(
      Listings = n(),
      N_Cities = n_distinct(.data[[city_col]], na.rm = TRUE),
      Cities   = paste(sort(unique(na.omit(.data[[city_col]]))), collapse = ", "),
      .groups  = "drop"
    ) %>%
    arrange(Listings)

  if (!is.null(min_listings)) {
    result <- result %>% filter(Listings < min_listings)
  }

  result
}

```

```{r dept-all}
cities_per_dept(df) %>%
  arrange(desc(Listings)) %>%
  kable(caption = "Cities per Department") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## ⚑ Cities with Fewer than 30 Listings

These cities have sparse representation and may need grouping or exclusion before modelling.

```{r city-sparse}
if (city_col %in% names(df)) {
  sparse_cities <- df %>%
    count(.data[[city_col]], name = "Listings") %>%
    filter(Listings < 30) %>%
    arrange(Listings)

  cat("Cities with fewer than 30 listings:", nrow(sparse_cities), "\n")
  cat("Total listings in those cities:", sum(sparse_cities$Listings), "\n")
  cat("City names:", paste(sparse_cities[[city_col]], collapse = ", "), "\n")

  sparse_cities %>%
    kable(caption = "Cities with fewer than 30 listings") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)

  ggplot(sparse_cities %>%
           mutate(!!city_col := fct_reorder(.data[[city_col]], Listings)),
         aes(x = .data[[city_col]], y = Listings, fill = Listings)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = Listings), hjust = -0.2, size = 3) +
    scale_fill_gradient(low = "#f7c59f", high = "#e05c1a") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
    coord_flip() +
    labs(title = "Cities with Fewer than 30 Listings",
         x = "City", y = "Listings") +
    theme_minimal(base_size = 11)
}
```

## Median Price per City (Top Cities)

```{r city-price}
if (city_col %in% names(df) && price_col %in% names(df)) {
  city_price <- df %>%
    filter(!is.na(.data[[price_col]]), !is.na(.data[[city_col]])) %>%
    group_by(.data[[city_col]]) %>%
    summarise(
      Median_Price = median(.data[[price_col]], na.rm = TRUE),
      Mean_Price   = mean(.data[[price_col]],   na.rm = TRUE),
      Listings     = n()
    ) %>%
    arrange(desc(Median_Price)) %>%
    slice_head(n = top_n_val) %>%
    mutate(!!city_col := fct_reorder(.data[[city_col]], Median_Price, .desc = TRUE))

  ggplot(city_price, aes(x = .data[[city_col]], y = Median_Price, fill = Median_Price)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = comma(round(Median_Price))), vjust = -0.4, size = 3) +
    scale_fill_gradient(low = "#a8c8e8", high = "#1a5fa8") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)), labels = comma) +
    labs(title = paste("Top", top_n_val, "Cities by Median Listing Price"),
         x = "City", y = "Median Price") +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```

---

# Department Analysis (Postal Code)

Departments are identified by the **first two digits** of `postal_code`. [web:16]

## Listings per Department

```{r dept-listings}
postal_col <- "postal_code"   # adjust if needed

if (postal_col %in% names(df)) {
  # Build department column
  df_dept <- df %>%
    mutate(department = str_sub(as.character(.data[[postal_col]]), 1, 2))

  dept_counts <- df_dept %>%
    count(department, name = "Listings") %>%
    arrange(desc(Listings)) %>%
    mutate(department = fct_reorder(department, Listings, .desc = TRUE))

  # ── Sparse departments (<30): names + their cities ──────────────────────────
  sparse_depts <- dept_counts %>%
    filter(Listings < 30) %>%
    arrange(Listings)

  cat("Departments with fewer than 30 listings:", nrow(sparse_depts), "\n")
  cat("Department codes:", paste(sparse_depts$department, collapse = ", "), "\n\n")

  if (nrow(sparse_depts) > 0) {
    # For each sparse department, list its cities
    sparse_dept_cities <- df_dept %>%
      filter(department %in% sparse_depts$department) %>%
      group_by(department) %>%
      summarise(
        Listings = n(),
        Cities   = paste(sort(unique(.data[[city_col]])), collapse = ", ")
      ) %>%
      arrange(Listings)

    cat("Cities per sparse department:\n")
    for (i in seq_len(nrow(sparse_dept_cities))) {
      cat(sprintf("  [%s] (%d listings): %s\n",
                  sparse_dept_cities$department[i],
                  sparse_dept_cities$Listings[i],
                  sparse_dept_cities$Cities[i]))
    }
    cat("\n")

    # Formatted table
    sparse_dept_cities %>%
      kable(caption = "Departments with fewer than 30 listings and their cities") %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
  }
  # ────────────────────────────────────────────────────────────────────────────

  # Full listing-count table for all departments
  dept_counts %>%
    kable(caption = "Listings per Department (first two digits of postal code)") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

  # Bar chart — all departments
  ggplot(dept_counts, aes(x = department, y = Listings, fill = Listings)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = comma(Listings)), vjust = -0.4, size = 3) +
    scale_fill_gradient(low = "#a8c8e8", high = "#1a5fa8") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)), labels = comma) +
    labs(title = "Number of Listings per Department",
         x = "Department Code", y = "Listings") +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

} else {
  cat("Column '", postal_col, "' not found. Update postal_col above.\n")
}
```

## Median Price per Department

```{r dept-price}
if (postal_col %in% names(df) && price_col %in% names(df)) {
  dept_price <- df %>%
    mutate(department = str_sub(as.character(.data[[postal_col]]), 1, 2)) %>%
    filter(!is.na(.data[[price_col]])) %>%
    group_by(department) %>%
    summarise(
      Median_Price = median(.data[[price_col]], na.rm = TRUE),
      Mean_Price   = mean(.data[[price_col]],   na.rm = TRUE),
      Listings     = n()
    ) %>%
    arrange(desc(Median_Price)) %>%
    mutate(department = fct_reorder(department, Median_Price, .desc = TRUE))

  ggplot(dept_price, aes(x = department, y = Median_Price, fill = Median_Price)) +
    geom_col(show.legend = FALSE) +
    geom_text(aes(label = comma(round(Median_Price))), vjust = -0.4, size = 3) +
    scale_fill_gradient(low = "#a8c8e8", high = "#1a5fa8") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1)), labels = comma) +
    labs(title = "Median Listing Price per Department",
         x = "Department Code", y = "Median Price") +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}
```
```{r region-analysis}
# ── Listings per region ───────────────────────────────────────────────────────
region_counts <- df %>%
  filter(region != "unknown") %>%
  count(region, name = "Listings") %>%
  arrange(desc(Listings)) %>%
  mutate(region = fct_reorder(region, Listings, .desc = TRUE))

ggplot(region_counts, aes(x = region, y = Listings, fill = Listings)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(Listings)), vjust = -0.4, size = 3) +
  scale_fill_gradient(low = "#a8c8e8", high = "#1a5fa8") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), labels = comma) +
  labs(title = "Number of Listings per Region", x = "Region", y = "Listings") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# ── Median price per region ───────────────────────────────────────────────────
region_price <- df %>%
  filter(region != "unknown", !is.na(price)) %>%
  group_by(region) %>%
  summarise(
    Median_Price = median(price, na.rm = TRUE),
    Mean_Price   = mean(price,   na.rm = TRUE),
    Listings     = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(Median_Price)) %>%
  mutate(region = fct_reorder(region, Median_Price, .desc = TRUE))

ggplot(region_price, aes(x = region, y = Median_Price, fill = Median_Price)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = comma(round(Median_Price))), vjust = -0.4, size = 3) +
  scale_fill_gradient(low = "#a8c8e8", high = "#1a5fa8") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)), labels = comma) +
  labs(title = "Median Listing Price per Region", x = "Region", y = "Median Price") +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r dept-sparse}
sparse <- cities_per_dept(df, min_listings = 30)

cat("Departments with fewer than 30 listings:", nrow(sparse), "\n")
cat("Department codes:", paste(sparse$department, collapse = ", "), "\n\n")

for (i in seq_len(nrow(sparse))) {
  cat(sprintf("  [%s] (%d listings, %d cities): %s\n",
              sparse$department[i],
              sparse$Listings[i],
              sparse$N_Cities[i],
              sparse$Cities[i]))
}

sparse %>%
  kable(caption = "Sparse Departments (< 30 listings) with their Cities") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  row_spec(seq_len(nrow(sparse)), background = "#fff3cd")
```

---

# Property Type Analysis

## ⚑ Property Types with Fewer than 30 Listings

```{r proptype-sparse}
# ── config ────────────────────────────────────────────────────────────────────
prop_col <- "property_type"   # adjust to your actual column name
# ─────────────────────────────────────────────────────────────────────────────

if (prop_col %in% names(df)) {
  proptype_counts <- df %>%
    count(.data[[prop_col]], name = "Listings") %>%
    arrange(Listings)

  sparse_proptypes <- proptype_counts %>% filter(Listings < 30)

  cat("Property types with fewer than 30 listings:", nrow(sparse_proptypes), "\n")
  cat("Property type names:", paste(sparse_proptypes[[prop_col]], collapse = ", "), "\n")

  proptype_counts %>%
    mutate(Sparse = ifelse(Listings < 30, "< 30", "≥ 30")) %>%
    kable(caption = "Listings per Property Type (flagged if < 30)") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE) %>%
    row_spec(which(proptype_counts$Listings < 30), background = "#fff3cd")

  if (nrow(sparse_proptypes) > 0) {
    ggplot(proptype_counts %>%
             mutate(!!prop_col := fct_reorder(.data[[prop_col]], Listings),
                    Sparse = ifelse(Listings < 30, "< 30", "≥ 30")),
           aes(x = .data[[prop_col]], y = Listings, fill = Sparse)) +
      geom_col() +
      geom_hline(yintercept = 30, linetype = "dashed",
                 colour = "#e05c1a", linewidth = 0.8) +
      geom_text(aes(label = Listings), hjust = -0.2, size = 3.2) +
      scale_fill_manual(values = c("< 30" = "#e05c1a", "≥ 30" = "#1a5fa8")) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
      coord_flip() +
      labs(title = "Listings per Property Type (dashed = 30-listing threshold)",
           x = "Property Type", y = "Listings", fill = "Count category") +
      theme_minimal(base_size = 12)
  }
} else {
  cat("Column '", prop_col, "' not found. Update prop_col above.\n")
}
```

---

# Price vs Variables

```{r fn-plot-price-vs-var}
# ══════════════════════════════════════════════════════════════════════════════
# plot_price_vs_var()
#
# Plots both price and log(price) against a single predictor variable.
# Automatically detects whether the predictor is numeric or categorical and
# selects the appropriate chart type:
#
#   Numeric predictor  → scatter plot with a loess smooth (price) and
#                        log10(price) on the y-axis for the second panel
#   Categorical predictor → boxplots of price and log10(price) per level,
#                           levels ordered by median price (descending),
#                           with n-labels on top of each box
#
# Args:
#   data         : data frame containing the data
#   var          : predictor column name (string)
#   price_col    : name of the price column (default "price")
#   top_n        : for categorical vars with many levels, show only the top_n
#                  levels by median price (NULL = show all)
#   point_alpha  : transparency for scatter points (default 0.25)
#   point_size   : size of scatter points (default 0.8)
#   color_raw    : fill/colour for the price panel (default "#1a5fa8")
#   color_log    : fill/colour for the log(price) panel (default "#6a3fa8")
#
# Returns: invisibly, a named list with elements $raw and $log (ggplot objects)
#
# Usage:
#   plot_price_vs_var(df, "surface")
#   plot_price_vs_var(df, "property_type", top_n = 15)
#   plot_price_vs_var(df, "rooms", color_raw = "#2ca02c")
# ══════════════════════════════════════════════════════════════════════════════

plot_price_vs_var <- function(data,
                              var,
                              price_col   = "price",
                              top_n       = NULL,
                              min_n       = 30,       # drop categorical levels with fewer listings
                              point_alpha = 0.25,
                              point_size  = 0.8,
                              color_raw   = "#1a5fa8",
                              color_log   = "#6a3fa8") {

  # ── Validate inputs ──────────────────────────────────────────────────────────
  missing_cols <- setdiff(c(var, price_col), names(data))
  if (length(missing_cols) > 0)
    stop("Column(s) not found in data: ", paste(missing_cols, collapse = ", "))

  # ── Working data: drop NA price and zero/negative prices for log scale ───────
  d <- data %>%
    filter(!is.na(.data[[price_col]]), .data[[price_col]] > 0,
           !is.na(.data[[var]]))

  is_numeric_var <- is.numeric(d[[var]])

  # ── NUMERIC predictor ────────────────────────────────────────────────────────
  if (is_numeric_var) {

    p_raw <- ggplot(d, aes(x = .data[[var]], y = .data[[price_col]])) +
      geom_point(alpha = point_alpha, size = point_size, colour = color_raw) +
      geom_smooth(method = "loess", formula = y ~ x,
                  colour = "#e05c1a", fill = "#f7c59f", linewidth = 0.9) +
      scale_y_continuous(labels = comma) +
      labs(
        title = paste("Price vs", var),
        x     = var,
        y     = "Price"
      ) +
      theme_minimal(base_size = 12)

    p_log <- ggplot(d, aes(x = .data[[var]], y = .data[[price_col]])) +
      geom_point(alpha = point_alpha, size = point_size, colour = color_log) +
      geom_smooth(method = "loess", formula = y ~ x,
                  colour = "#e05c1a", fill = "#f7c59f", linewidth = 0.9) +
      scale_y_log10(labels = comma) +
      labs(
        title = paste("log10(Price) vs", var),
        x     = var,
        y     = "log10(Price)"
      ) +
      theme_minimal(base_size = 12)

  # ── CATEGORICAL predictor ────────────────────────────────────────────────────
  } else {

    # Order levels by median price; drop sparse levels; optionally keep only top_n
    level_order <- d %>%
      group_by(.data[[var]]) %>%
      summarise(med = median(.data[[price_col]], na.rm = TRUE),
                n   = n(),
                .groups = "drop") %>%
      filter(n >= min_n) %>%          # ← drop levels below the min_n threshold
      arrange(desc(med))

    if (nrow(level_order) == 0) {
      message("plot_price_vs_var: no levels of '", var,
              "' have >= ", min_n, " listings — skipping.")
      return(invisible(NULL))
    }

    if (!is.null(top_n)) level_order <- slice_head(level_order, n = top_n)

    d <- d %>% filter(.data[[var]] %in% level_order[[var]]) %>%
      mutate(!!var := factor(.data[[var]], levels = level_order[[var]]))

    # n-labels (counts per level)
    n_labels <- d %>%
      count(.data[[var]]) %>%
      rename(n_count = n)

    # Build subtitle noting any dropped sparse levels
    n_dropped  <- n_distinct(d[[var]]) - nrow(level_order)
    sparse_note <- if (n_dropped > 0)
      paste0(" [", n_dropped, " level(s) with <", min_n, " listings hidden]") else ""

    p_raw <- ggplot(d, aes(x = .data[[var]], y = .data[[price_col]])) +
      geom_boxplot(fill = color_raw, colour = "#333333",
                   outlier.colour = "#e05c1a", outlier.alpha = 0.3,
                   outlier.size = 0.7) +
      geom_text(data = n_labels,
                aes(x = .data[[var]], y = Inf,
                    label = paste0("n=", comma(n_count))),
                vjust = 1.4, size = 2.8, inherit.aes = FALSE) +
      scale_y_continuous(labels = comma) +
      labs(
        title = paste0("Price by ", var,
                       if (!is.null(top_n)) paste0(" (top ", top_n, " levels)") else "",
                       sparse_note),
        x = var,
        y = "Price"
      ) +
      theme_minimal(base_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))

    p_log <- ggplot(d, aes(x = .data[[var]], y = .data[[price_col]])) +
      geom_boxplot(fill = color_log, colour = "#333333",
                   outlier.colour = "#e05c1a", outlier.alpha = 0.3,
                   outlier.size = 0.7) +
      geom_text(data = n_labels,
                aes(x = .data[[var]], y = Inf,
                    label = paste0("n=", comma(n_count))),
                vjust = 1.4, size = 2.8, inherit.aes = FALSE) +
      scale_y_log10(labels = comma) +
      labs(
        title = paste0("log10(Price) by ", var,
                       if (!is.null(top_n)) paste0(" (top ", top_n, " levels)") else "",
                       sparse_note),
        x = var,
        y = "log10(Price)"
      ) +
      theme_minimal(base_size = 12) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
  }

  print(p_raw)
  print(p_log)
  invisible(list(raw = p_raw, log = p_log))
}
```

## Price vs All Variables

```{r price-vs-all-vars, fig.height=5}
# ── Config ────────────────────────────────────────────────────────────────────
# Variables to skip (the target itself and any IDs / high-cardinality text)
skip_vars <- c(price_col, "id_annonce", "postal_code", "department", "region")

# For categorical variables with many levels, show only the top N by median price
cat_top_n <- 20

# Categorical levels with fewer than this many listings are excluded from plots
cat_min_n <- 30
# ─────────────────────────────────────────────────────────────────────────────

vars_to_plot <- setdiff(names(df), skip_vars)

for (v in vars_to_plot) {
  cat("\n\n---\n\n## Price vs `", v, "`\n\n", sep = "")

  # Determine top_n: apply only for categorical with > cat_top_n distinct levels
  is_cat  <- is.character(df[[v]]) || is.factor(df[[v]])
  n_lvls  <- n_distinct(df[[v]], na.rm = TRUE)
  use_top <- if (is_cat && n_lvls > cat_top_n) cat_top_n else NULL

  plot_price_vs_var(df, var = v, price_col = price_col,
                    top_n = use_top, min_n = cat_min_n)
}
```

---

```{r corrplot, fig.width=10, fig.height=10}
# Encode department and region as their median price (numeric proxy for correlation)
dept_median  <- df %>% group_by(department) %>%
  summarise(dept_median_price  = median(price, na.rm = TRUE), .groups = "drop")
region_median <- df %>% group_by(region) %>%
  summarise(region_median_price = median(price, na.rm = TRUE), .groups = "drop")

df_cor <- df %>%
  left_join(dept_median,   by = "department") %>%
  left_join(region_median, by = "region")     %>%
  select(where(is.numeric))   # now includes dept_median_price + region_median_price

cor_matrix <- cor(df_cor, use = "complete.obs")

corrplot(
  cor_matrix,
  method      = "color",
  type        = "upper",
  tl.cex      = 0.7,
  tl.col      = "black",
  addCoef.col = "black",
  number.cex  = 0.5,
  title       = "Correlation Matrix – Numeric Variables (incl. geo medians)",
  mar         = c(0, 0, 1.5, 0)
)
```
---

# Session Info

```{r session-info}
sessionInfo()
```
